Beam-Helicity and Beam-Charge Asymmetries \\ in the \red{DVCS and Bethe--Heitler processes} on \\ an Unpolarised Proton Target \underline{Paper Tag: PDVCS0607 (v2.3)} \underline{Drafting Committee: DC90 (\today)}

### 2012-01-18 results_dual.tex
Updated with all comments from DC meeting. Need to fix plots and order of subscripts.

\section{Results}
The results for the beam-helicity and beam-charge asymmetry amplitudes extracted from the 2006-2007 hydrogen data sets in this work and results extracted in Ref.~\cite{Air09} from the previously published 1996-2005 data set are compared in figures~\ref{release_bsa_0607} and~\ref{release_bca_0607}. Each of the asymmetry amplitudes is shown extracted in one bin over all kinematic variables (``overall'') and also projected against $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The beam-helicity asymmetry amplitudes are subject to an additional scale uncertainty from the measurement of the beam polarisation, which is stated in the captions of the figures.
\begin{figure}
\begin{center}
\includegraphics[width=15cm,keepaspectratio]{macros/bsa_both_morgan}
  \caption{Beam-helicity asymmetry amplitudes extracted separately from
the 1996-2005 (open triangles)~\cite{Air09} and 2006-2007 (filled squares)
hydrogen data. The error bars represent the statistical uncertainties. The error bands represent the systematic uncertainties.
An additional 2.8\,\% and 3.4\,\% scale uncertainty for the 1996-2005 and
2006-2007 data respectively is present in the amplitudes due to the uncertainty of
the beam polarisation measurement. The simulated fractional contribution from associated production to the yield in each kinematic bin is shown in the bottom row.}
 \label{release_bsa_0607}
\end{center}
 \end{figure}

A statistical test (Student's test) was applied in order to check for possible incompatibility between the asymmetry amplitudes extracted from the two data sets. Only the statistical uncertainties were employed in this test as the largest contributions to the systematic uncertainties are largely correlated, i.e. effects from detector resolution, acceptance and smearing. This test revealed no significant evidence for incompatibility between the data sets. The beam-helicity and beam-charge asymmetry amplitudes can therefore be extracted from the entire hydrogen data set recorded during the entire experimental operation of H{\sc ermes}.
\begin{figure}
\begin{center}
 \includegraphics[width=15cm,keepaspectratio]{macros/bca_both_morgan}
  \caption{Beam-charge asymmetry amplitudes extracted separately from the 1996-2005 (open triangles)~\cite{Air09} and 2006-2007 (filled squares) hydrogen data.
The error bars represent the statistical uncertainties. The error bands represent the systematic uncertainties. The simulated fractional contribution from associated production to the yield in each kinematic bin is shown in the bottom row.}
 \label{release_bca_0607}
\end{center}
 \end{figure}

@@

The results of the beam-helicity and beam-charge asymmetry amplitudes extracted from the complete 1996-2007 hydrogen sample are shown in figures~\ref{bsa_xbjrange} and \ref{bca_xbjrange}. The number of analysable events available from the 2006-2007 data set (70352) is approximately three times greater than the number of events recorded in the 1996-2005 sample (24817). The  asymmetry amplitudes extracted from the complete 1996-2007 data set thus resemble the 2006-2007 result. This resemblance is not so evident for the beam-helicity asymmetry amplitudes as it is for the beam-charge asymmetry amplitudes, because the beam polarisation was lower in 2006 and 2007 than in 1996-2005 and thus the 2006-2007 data has a lower weighting in the combined fit. The \red{major contributions to the} systematic error bands associated with the asymmetry amplitudes extracted from the combined data set were determined using Monte Carlo simulations as explained in section~4\red{, i.e. contributions from acceptance, smearing, finite bin widths and misalignment. \red{The background in the data sample is estimated using the method from refs.~\cite{Zei09, Bur10}.} The missing-mass uncertainty for the combined data \red{sets were} calculated using the procedure described in that section~4 and the results were \red{averaged}. The total systematic uncertainties for the combined results are the result of these three independent contributions added in quadrature.}

@@

The $A_{\textrm{C}}^{\cos(n\phi)}$ amplitudes are shown in figure~\ref{bca_xbjrange}. The systematic uncertainties are estimated \red{using} the same procedure as \red{was} used to estimate th\red{ose for} the beam-helicity asymmetries. The leading twist $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ amplitudes are both non-zero. There is an expected relationship between $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ as they depend on the same $\mathcal{C}$-function. The kinematic suppression of $A_{\textrm{C}}^{\cos(0\phi)}$ with regards to $A_{\textrm{C}}^{\cos\phi}$ \red{is approximately fulfilled}. The measured amplitudes are found to diverge with opposite sign from zero at increasing values of $-t$ and they indicate a weak increase with $x_{\textrm{B}}$ and $Q^{2}$. The $A_{\textrm{C}}^{\cos(2\phi)}$ and $A_{\textrm{C}}^{\cos(3\phi)}$ amplitudes are both consistent with zero over the \red{whole} range in $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The $A_{C}^{\cos(2\phi)}$ amplitude is related to twist-3 GPDs and $A_{\textrm{C}}^{\cos(3\phi)}$ relates to gluon helicity-flip GPDs. Both of these amplitudes are expected to be suppressed at H{\sc ermes} kinematic conditions compared to the leading twist amplitudes.

@@

The dash-dotted curves in figure~\ref{bsa_xbjrange} shows the result of a calculation from a fit based on a quark-diquark model with a Regge-inspired term that is included in order to describe accurately parton distribution functions at low $x$ values~\cite{Liu11}. The ``Regge'' term is extended to include contributions that determine the $t$-dependence of the corresponding GPD. The model incorporates fits to global deep-inelastic and elastic scattering data (to account for the $\xi$-independent limits and moments of the underlying GPDs) and DVCS data from Jefferson Laboratory (to describe the skewness dependence). The model describes the $t$-projections of the $A^{\sin\phi}_{\textrm{LU}}$ amplitude reported here well, but the projections in the other kinematic variables are not as well described. The model describes the trends of the $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ asymmetry amplitudes well.

In order to provide more detailed information that can be used in future fits, in particular for the determination of the entanglement of the skewness and $-t$ dependences of GPDs, the amplitudes already presented in Figs.~\ref{bsa_xbjrange} and~\ref{bca_xbjrange} are shown as a function of $-t$ for three different ranges of $x_{\textrm{B}}$ in Figs.~\ref{bsa_xbjrange2} and~\ref{bca_xbjrange2}. These figures represent the kinematic dependences of the amplitudes in a less-correlated manner than the one dimensional projections\red{: within experimental uncertainty, there is no evidence of a correlation between the $-t$ dependence  and the $x_{\textrm{B}}$ dependence for any of the amplitudes}. The results from this paper will be made available in the Durham Database in the same 4 bin format as used in previous analyses at H{\sc ermes}~\cite{Air08,Air09,Air10b}.

>>>>

\section{Results}
\red{In figures~\ref{release_bsa_0607} and~\ref{release_bca_0607}, results} for the beam-helicity and beam-charge asymmetry amplitudes extracted from the 2006-2007 hydrogen data sets in this work \red{are compared with} results extracted in  from the 1996-2005 data set \red{published previously~\cite{Air09}}. Each of the asymmetry amplitudes is shown extracted in one bin over all kinematic variables (``overall'') and also projected against $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The beam-helicity asymmetry amplitudes are subject to an additional scale uncertainty from the measurement of the beam polarisation, which is stated in the captions of the figures.
\begin{figure}
\begin{center}
\includegraphics[width=15cm,keepaspectratio]{macros/bsa_both_morgan}
  \caption{Beam-helicity asymmetry amplitudes extracted separately from
the 1996-2005 (open triangles)~\cite{Air09} and 2006-2007 (filled squares)
hydrogen data. The error bars represent the statistical uncertainties. The error bands represent the systematic uncertainties.
An additional 2.8\,\% and 3.4\,\% scale uncertainty for the 1996-2005 and
2006-2007 data respectively is present in the amplitudes due to the uncertainty of
the beam polarisation measurement. The simulated fractional contribution from associated production to the yield in each kinematic bin is shown in the bottom row.}
 \label{release_bsa_0607}
\end{center}
 \end{figure}

A statistical test (Student's \red{t-}test) was applied in order to check for possible incompatibility between the asymmetry amplitudes extracted from the two data sets. Only the statistical uncertainties were employed in this test as the largest contributions to the systematic uncertainties\red{, i.e. effects from detector resolution, acceptance, misalignment and smearing, are largely correlated}. This test revealed no significant evidence for incompatibility between the data sets. The beam-helicity and beam-charge asymmetry amplitudes can therefore be extracted from the entire hydrogen data set recorded during the entire experimental operation of H{\sc ermes}.
\begin{figure}
\begin{center}
 \includegraphics[width=15cm,keepaspectratio]{macros/bca_both_morgan}
  \caption{Beam-charge asymmetry amplitudes extracted separately from the 1996-2005 (open triangles)~\cite{Air09} and 2006-2007 (filled squares) hydrogen data.
The error bars represent the statistical uncertainties. The error bands represent the systematic uncertainties. The simulated fractional contribution from associated production to the yield in each kinematic bin is shown in the bottom row.}
 \label{release_bca_0607}
\end{center}
 \end{figure}

@@

The results of the beam-helicity and beam-charge asymmetry amplitudes extracted from the complete 1996-2007 hydrogen sample are shown in figures~\ref{bsa_xbjrange} and \ref{bca_xbjrange}. The number of analysable events available from the 2006-2007 data set (70352) is approximately three times greater than the number of events recorded in the 1996-2005 sample (24817). The  asymmetry amplitudes extracted from the complete 1996-2007 data set thus resemble the 2006-2007 result. This resemblance is not so evident for the beam-helicity asymmetry amplitudes as it is for the beam-charge asymmetry amplitudes, because the beam polarisation was lower in 2006 and 2007 than in 1996-2005 and thus the 2006-2007 data has a lower weighting in the combined fit. The \red{major contributions to the} systematic error bands associated with the asymmetry amplitudes extracted from the combined data set were determined using Monte Carlo simulations as explained in section~4\red{, i.e. contributions from acceptance, smearing, finite bin widths and misalignment. \red{The background in the data sample is estimated using the method from refs.~\cite{Zei09, Bur10}.} The \red{uncertainty contributions due to the observed shifts in the missing-mass distributions} for the combined data \red{sets were} calculated using the procedure described in section~4 and the results were \red{averaged}. The total systematic uncertainties for the combined results are the result of these three independent contributions added in quadrature.}

@@

The $A_{\textrm{C}}^{\cos(n\phi)}$ amplitudes are shown in figure~\ref{bca_xbjrange}. The systematic uncertainties are estimated \red{using} the same procedure as \red{was} used to estimate th\red{ose for} the beam-helicity asymmetries. The leading twist $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ amplitudes are both non-zero. There is an expected relationship between $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ as they depend on the same $\mathcal{C}$-function. The kinematic suppression of $A_{\textrm{C}}^{\cos(0\phi)}$ with regard to $A_{\textrm{C}}^{\cos\phi}$ \red{is approximately fulfilled}. The measured amplitudes are found to diverge with opposite sign from zero at increasing values of $-t$ and they indicate a weak increase with $x_{\textrm{B}}$ and $Q^{2}$. The $A_{\textrm{C}}^{\cos(2\phi)}$ and $A_{\textrm{C}}^{\cos(3\phi)}$ amplitudes are both consistent with zero over the \red{whole} range in $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The $A_{C}^{\cos(2\phi)}$ amplitude is related to twist-3 GPDs and $A_{\textrm{C}}^{\cos(3\phi)}$ relates to gluon helicity-flip GPDs. Both of these amplitudes are expected to be suppressed at H{\sc ermes} kinematic conditions compared to the leading twist amplitudes.

@@

The dash-dotted curves in figure~\ref{bsa_xbjrange} shows the result of a calculation from a fit based on a quark-diquark model with a Regge-inspired term that is included in order to describe accurately parton distribution functions at low $x$ values~\cite{Liu11}. The ``Regge'' term is extended to include contributions that determine the $t$-dependence of the corresponding GPD. The model incorporates fits to global deep-inelastic and elastic scattering data (to account for the $\xi$-independent limits and moments of the underlying GPDs) and DVCS data from Jefferson Lab\red{.} (to describe the skewness dependence). The model describes the $t$-projections of the $A^{\sin\phi}_{\textrm{LU}}$ amplitude reported here well, but the projections in the other kinematic variables are not as well described. The model describes the trends of the $A_{\textrm{C}}^{\cos(0\phi)}$ and $A_{\textrm{C}}^{\cos\phi}$ asymmetry amplitudes well.

In order to provide more detailed information that can be used in future fits, in particular for the determination of the entanglement of the skewness and $-t$ dependences of GPDs, the amplitudes already presented in Figs.~\ref{bsa_xbjrange} and~\ref{bca_xbjrange} are shown as a function of $-t$ for three different ranges of $x_{\textrm{B}}$ in Figs.~\ref{bsa_xbjrange2} and~\ref{bca_xbjrange2}. These figures represent the kinematic dependences of the amplitudes in a less-correlated manner than the one dimensional projections\red{: within experimental uncertainty, there is no evidence of a correlation between the $-t$ dependence  and the $x_{\textrm{B}}$ dependence for any of the amplitudes}. The results from this paper will be made available in the Durham Database\red{. The results will also be made} available in the same 4 bin format as used in previous analyses at H{\sc ermes}~\cite{Air08,Air09,Air10b}.

### 2012-01-18 summary.tex
Updated with all comments from DC meeting. Need to fix plots and order of subscripts.

Beam-helicity and beam-charge asymmetries in the azimuthal distribution of real photons from leptoproduction on an unpolarised hydrogen target have been presented. These asymmetries were extracted from a\red{n} unpolarised hydrogen data set taken during the 2006 and 2007 operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were extracted previously from hydrogen data obtained during the 1996-2005 experimental period as described in ref.~\cite{Air09}. A comparison of the amplitudes extracted from these independent data sets has shown that they are compatible and the asymmetry amplitudes can therefore be extracted from the complete 1996-2007 event sample. The asymmetry amplitudes extracted from the combined data set are the most statistically precise DVCS measurements presented by H{\sc ermes}. There is a strong signal in the first harmonic of the interference contribution to the beam-helicity asymmetry. There are non-zero amplitudes in the zeroth and first harmonics of the beam-charge asymmetry. Each of the higher order asymmetry amplitudes is consistent with zero. The combined results are compared to calculations from ongoing work to fit GPD models to experimental data. All data amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

>>>>

Beam-helicity and beam-charge asymmetries in the azimuthal distribution of real photons from leptoproduction on an unpolarised hydrogen target have been presented. These asymmetries were extracted from a\red{n} unpolarised hydrogen data set taken during the 2006 and 2007 operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were extracted previously from hydrogen data obtained during the 1996-2005 experimental period as described in ref.~\cite{Air09}. A comparison of the amplitudes extracted from these independent data sets has shown that they are compatible and the asymmetry amplitudes can therefore be extracted from the complete 1996-2007 \red{data set}. The asymmetry amplitudes extracted from the com\red{plete} data set are the most statistically precise DVCS measurements presented by H{\sc ermes}. There is a strong signal in the first harmonic of the interference contribution to the beam-helicity asymmetry. There are non-zero amplitudes in the zeroth and first harmonics of the beam-charge asymmetry. Each of the higher order asymmetry amplitudes is consistent with zero. The combined results are compared to calculations from ongoing work to fit GPD models to experimental data. All data amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.
